\documentclass{amsart}
\usepackage{macros}
\title{Commentary on Bourbaki's ``Description of Formal Mathematics''}
\author{Alex Nelson}
\email{pqnelson@gmail.com}
\date{January 19, 2024}
\setcounter{section}{-1}
\begin{document}
\maketitle

\tableofcontents

\vfill\eject
\section{Preface}

\subsection{Urtext}
This is my commentary on the first chapter of Bourbaki's \textit{Theory of Sets}~\cite{bourbaki1970sets},
specifically the first chapter. I am mostly relying on the English
translation --- apparently the 1968 English translation of the French 1970
edition. Apparently the English translators had access to a time-machine.
(I suspect the publisher erred in the dates, but it's the softcover
English translation available right now by Springer.)

There are only one or two times I return to the French edition being
translated. The English translation appears to be faithful to the point
of being a literal translation. However, there is a critical passage
which omits an important phrase, and fails to capture the idiomatic
meaning of the passage in the original.

%% A lot of it is based on the
%% anachronistic perspective of trying to articulate Bourbaki's formalism
%% into something which could become a proof assistant. Most of the time I
%% will rely on the English translation, only rarely do I return to the
%% French edition and note differences in translations.

\subsection{Organization of these notes}
The ``sections'' and ``subsections'' are organized after Bourbaki's
sections and subsections. I make a number of observations, which are
organized into ``subsubsections''. So note ``1.3.4'' refers to section
1, subsection 3, and it's my fourth note on that section.

\subsubsection{Puzzles} There are a number of puzzles which are numbered
sequentially. These are questions for myself, which are really points of
discussion for the reader as well.

\subsection{Pronouns}
Bourbaki is both a group and an individual. I'm not sure what pronoun to
use to refer to Bourbaki, therefore I will switch between ``he/him'' and
``they/them'' according to what sounds good to my tone-deaf ear.

\subsection{Exegesis/Eisegesis}
I'm not trying to ``read in what I want'', but there are times when it
is too ambiguous for clearly making sense of what Bourbaki ``really meant''.
Therefore, I cannot claim to be clarifying Bourbaki's underlying intent.

But I knowingly deviate from Bourbaki's notation in the following ways:
Bourbaki uses Polish notation (``lispy syntax'') and then immediately
discards it, I just avoid it; Bourbaki uses a 2-dimensional syntax for
linkages, I just intuitively think of using de Bruijn indices.

\subsubsection{Anachronisms}
Bourbaki didn't actually adopt Hilbert's $\varepsilon$-calculus until
July 1950, when Chevellay wrote the fifth draft of the chapter on formal
mathematics~\cite{archives-no139}. The previous four drafts appear to be
written \emph{before} World War 2, so there's a span of more than a
decade separating these drafts. I mention this because a lot of the
concepts which Bourbaki introduces are rather clunky, and predate the
conventional ideas of a formal grammar (introduced by Chomsky in 1956~\cite{chomsky1956})
and compiler technology (introduced around \textit{c}.1954, but not popularized
until well into the 1960s).

\subsubsection{Methodology}
Raymond Chandler once remarked something along the lines of, ``People
who write about writing know nothing about how to write well.'' I fear
the same could be said about writing commentaries. Roughly, I am guided
by a desire to write a proof assistant. Consequently, I am interested in
forming a sort of ``rational reconstruction'' of Bourbaki's formal system.

Loosely, what I mean by a ``rational reconstruction'' is that I am
trying to implement these ideas in a computer program, so how can I
faithfully implement this? A lot of the ideas ``sound'' very close to
ideas that emerged later (formal grammars, production rules, judgements,
inference rules, inductive definitions, etc.). For me, a ``rational reconstruction''
is a \emph{translation} of the original text into a different formal
language which tries to preserve the critical aspects of the original
text. Because ``translation is interpretation'', this forces us to
interpret the original text and weigh what is important.

There is one or two instances where I have a ``close reading'' of
certain key passages of Bourbaki, consulting the French original. I'd
like to thank Stanley Oropesa for his help with certain French phrases. 

\subsection{Other commentaries}
Wayne Aitken~\cite{aitken2023} has written a good commentary about the
first chapter of Bourbaki~\cite{bourbaki1970sets}. Aitken has similar
goals as me, but wishes to clarify Bourbaki's rushed chapter (which
really is rather austere as a text).

Adrian Mathias has written a number of articles~\cite{mathias1992ignorance,mathias2002term,mathias2014hilbert}
about Bourbaki's formal system and its flaws. Mathias's articles are
classics and worth reading in their own right.

\section{Terms and Relations}

\input{1-1-signs-assemblies}
\input{1-2-substitution}
\input{1-3-formative-construction}
\input{1-4-formative-criteria}

\section{Theorems}
\input{2-1-axioms}
\input{2-2-proofs}
\input{2-3-substitutions-in-theory}
\input{2-4-comparing-theories}

\section{Logical Theories}
\input{3-1-axioms}
\input{3-2-first-consequences}
\input{3-3-methods-of-proof}
\input{3-4-conjunction}
\input{3-5-equivalence}

\section{Quantified Theories}
\input{4-1-quantifiers}
\input{4-2-axioms}
\input{4-3-properties}
\input{4-4-typical}

\section{Equalitarian Theories}
\input{5-1-axioms}

\begin{thebibliography}{99}
\bibitem{abadi1991explicit}
  M. Abadi, L. Cardelli, P-L. Curien and J-J. Levy,
  ``Explicit Substitutions''.
  \textit{Journal of Functional Programming} \textbf{1} no.4 (1991) 375--416.
\bibitem{aitken2023} Wayne Aitken,
  ``Bourbaki, Theory of Sets, Chapter I, \textit{Description of Formal Mathematics}: Summary and Commentary''.
  Manuscript dated June 2022,
  \url{https://public.csusm.edu/aitken_html/Essays/Bourbaki/BourbakiSetTheory1.pdf}
\bibitem{mizar} Grzegorz Bancerek, Czesław Byli\'{n}ski, Adam Grabowski, Artur Korni\l{}owicz, Roman Matuszewski, Adam Naumowicz, Karol Pak, and Josef Urban.
  ``Mizar: State-of-the-art and beyond.''
  In Manfred Kerber, Jacques Carette, Cezary Kaliszyk, Florian Rabe, and Volker Sorge, editors, \textit{Intelligent Computer Mathematics}, volume 9150 of Lecture Notes in Computer Science, pages 261--279. Springer International Publishing, 2015. \texttt{doi:\href{http://dx.doi.org/10.1007/978-3-319-20615-8_17}{10.1007/978-3-319-20615-8 17}}.
\bibitem{bernays1926} Paul Bernays, ``Axiomatische Untersuchungen des
Aussagen-Kalkuls der \textit{Principia Mathematica}.''
\textit{Mathematische Zeitschrift} \textbf{25} (1926) 305--320;
translated into English in Richard Zach's \textit{Universal Logic: An
  Anthology} (2012) pp.43--58.
\bibitem{bourbaki1970sets} Nicolas Bourbaki,
  \textit{Theory of Sets}.
  Springer, softcover reprint, 1968 English translation.
\bibitem{archives-no139}
  Claude Chevalley, ``Editorial No. 139. Sets. Chapter I (state 5).''
  Manuscript with handwritten note dating it as July 1950.
  Identifier \verb#R139_nbr042#, accessed January 17, 2024 \url{http://archives-bourbaki.ahp-numerique.fr/items/show/548}
  --- first introduced the Hilbert $\varepsilon$ operator into Bourbaki's system.
\bibitem{chomsky1956}
  Noam Chomsky,
  ``Three models for the description of language''.
  \textit{IRE Transactions on Information Theory} \textbf{2} no.3 (1956) 113--124.
\bibitem{corcoran2006schemata}
  John Corcoran,
  ``Schemata: The Concept of Schema in the History of Logic''.
  \textit{Bulletin of Symbolic Logic} \textbf{12} no.2 (2006) 219--240;
  \texttt{doi:\href{https://doi.org/10.2178/bsl/1146620060}{10.2178/bsl/1146620060}}
\bibitem{mathias1992ignorance}
  Adrian RD Mathias,
  ``The ignorance of Bourbaki''.
  \textit{Mathematical Intelligencer} \textbf{14} no.3 (1992) 4--13
\bibitem{mathias2002term}
  Adrian RD Mathias,
  ``A term of length 4,523,659,424,929''.
  \textit{Synthese} \textbf{133} (2002) 75--86
\bibitem{mathias2014hilbert}
  Adrian RD Mathias,
  ``Hilbert, Bourbaki and the scorning of logic''.
  In \textit{Infinity and truth} (World Scientific, 2014) pp.47--156.
  \url{https://www.dpmms.cam.ac.uk/~ardm/lbmkemily5.pdf}
\bibitem{wiedijk2000vernacular}
  Freek Wiedijk,
  ``The Mathematical Vernacular''.
  Unpublished manuscript, 2000
  \url{http://www.cs.ru.nl/F.Wiedijk/notes/mv.pdf}
\end{thebibliography}
\end{document}
